---
title: "p8105_hw5_cz2955"
output: html_document
name: Chenhai Zheng
date: "2025-11-12"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(broom)
library(purrr)
knitr::opts_chunk$set(echo = TRUE)
```

## Question1
### Function
```{r}
same_birth<-function(n){
birthdays = sample(1:365, n, replace = TRUE)
  any(duplicated(birthdays))
}
```
### Simulation
```{r}
set.seed(36501)  

n_values <- 2:50 
simulations <- 10000 
probabilities <- sapply(n_values, function(n) {
  mean(replicate(simulations, same_birth(n)))
})

```
### Plot
```{r}
plot(n_values, probabilities, type = "b", pch = 19, col = "blue",
     xlab = "Group Size (n)",
     ylab = "Probability of Shared Birthday",
     main = "Birthday Paradox Simulation")

abline(h = 0.5, col = "red", lty = 2)

```

## Question2
### Data
```{r}
set.seed(36501)

n <- 30
sigma <- 5
mu_values <- 0:6
n_sim <- 5000
simulate_mu <- function(mu) {
  replicate(n_sim, {
    x <- rnorm(n, mean = mu, sd = sigma)
    test <- t.test(x, mu = 0)
    tibble(
      mu_true = mu,
      mu_hat = mean(x),
      p_value = test$p.value
    )
  }, simplify = FALSE) |> bind_rows()
}

results <- map_df(mu_values, simulate_mu)

results <- results %>%
  mutate(reject = p_value < 0.05)
results 
```

### Plot1
```{r}
power_summary <- results %>%
  group_by(mu_true) %>%
  summarize(power = mean(reject))

ggplot(power_summary, aes(x = mu_true, y = power)) +
  geom_line(color = "blue", linewidth = 1.2) +
  geom_point(size = 2) +
  labs(
    x = "True mean (μ)",
    y = "Proportion of rejections (Power)",
    title = "Power of one-sample t-test vs Effect Size"
  ) +
  theme_minimal()


```

### plot2
```{r}
 mu_summary <- results %>%
  group_by(mu_true) %>%
  summarize(
    avg_mu_hat_all = mean(mu_hat),
    avg_mu_hat_reject = mean(mu_hat[reject])
  )

ggplot(mu_summary, aes(x = mu_true)) +
  geom_line(aes(y = avg_mu_hat_all, color = "All samples"), linewidth = 1.2) +
  geom_point(aes(y = avg_mu_hat_all, color = "All samples")) +
  geom_line(aes(y = avg_mu_hat_reject, color = "Rejected only"), linewidth = 1.2, linetype = "dashed") +
  geom_point(aes(y = avg_mu_hat_reject, color = "Rejected only")) +
  labs(
    x = "True mean (μ)",
    y = "Average estimate (μ̂)",
    title = "Average estimates across all samples vs rejected samples",
    color = ""
  ) +
  theme_minimal()

```
Ans：No, the sample average of $\hat\mu $ across tests for which the null was rejected is not approximately equal to the true value of $\mu$, especially when the true effect size is small.

This happens because we are conditioning on rejection of the null hypothesis (i.e.,𝑝<0.05). When the true mean $\mu$ is close to 0, only those samples that happened — by random chance — to produce unusually large sample means will yield significant results. As a result, the average of $\hat\mu $ among the “significant” samples is biased upward relative to the true $\mu $.

As the true effect size increases and the test’s power approaches 1, this selection bias disappears — almost all samples are significant, and the conditional average of $\hat\mu $ converges to the true value.

## Question3
### Baltimore Data
```{r}
homicide_data <- read_csv("homicide_data.csv")
homicide_data <- homicide_data %>%
  mutate(city_state = str_c(state, city, sep = ", "))%>%
  mutate(
    unsolved = if_else(
      disposition %in% c("Closed without arrest", "Open/No arrest"),
      1, 0
    )
  )
head(homicide_data)

city_summary <- homicide_data %>%
  group_by(city_state) %>%
  summarize(
    total = n(),
    unsolved = sum(unsolved),
    .groups = "drop"
  )

baltimore <- city_summary %>%
  filter(city_state == "MD, Baltimore")

baltimore_test <- prop.test(
  x = baltimore$unsolved,
  n = baltimore$total
)

baltimore_result <- tidy(baltimore_test)
baltimore_result

```
### Every City
```{r}
city_results <- city_summary %>%
  mutate(
    test = map2(unsolved, total, ~ prop.test(.x, .y) %>% tidy())
  ) %>%
  unnest(test) %>%
  select(city_state, estimate, conf.low, conf.high, total) %>%
  arrange(desc(estimate))
city_results
```

### Plot
```{r}
ggplot(city_results, aes(
  x = reorder(city_state, estimate),
  y = estimate
)) +
  geom_point(color = "darkorange", size = 2.2) +
  geom_errorbar(
    aes(ymin = conf.low, ymax = conf.high),
    width = 0.25,
    color = "gray40"
  ) +
  coord_flip() +
  labs(
    title = "Proportion of Unsolved Homicides by City",
    x = "City",
    y = "Estimated Proportion (95% CI)"
  ) +
  theme_minimal(base_size = 12) +         
  theme(axis.text.y = element_text(size = 8)) 


```






